---
title: "machine learning"
author: "kad!"
date: "7/2/2020"
output: html_document
---

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Preprocessing

```{r, cache=TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

## Downloading the Data

```{r, cache = TRUE}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile)
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile)
}
```  

## Reading the Data
 
```{r, cache = TRUE}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
The trainRaw contains 19622 observations and 160 variables, while the testRaw contains 20 observations and 160 variables.

## Clean the data

```{r, cache = TRUE}
sum(complete.cases(trainRaw))
```

we are going to remove the columns that contain NA missing values.

```{r, cache = TRUE}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  

we get rid of some columns that do not contribute much to the accelerometer 
measurements.

```{r, cache = TRUE}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

The cleaned trainCleaned contains 19622 observations and 53 variables. The testCleaned contains 20 observations and 53 variables.

## Slicing the data

We split the cleaned training set into a pure training data set (70%) and a validation data set (30%).

```{r, cache = TRUE}

set.seed(22222) 

inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling
  
```{r, cache = TRUE}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
  
```{r, cache = TRUE}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```

```{r, cache = TRUE}
accuracy <- postResample(predictRf, testData$classe)
accuracy
outofsampleerror <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
outofsampleerror
```
So, the estimated accuracy of the model is 99.37% and the estimated out of sample error is 0.66%.

## Predicting for Test Data Set
  
```{r, cache = TRUE}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

## Figures

Correlation Matrix Visualization  

```{r, cache = TRUE}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```

Decision Tree Visualization

```{r, cache = TRUE}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel)
```
